High-Performance Data Analysis on JANUS using Apache Spark

### 2013-12-18 paper.tex
more revisions to PageRank description

\subsection{Testing}
Following the example from Google's first paper on MapReduce, the first test we
performed was a parallel grep \citep{dean-mapreduce}. Here an application would
scan through a large test file, looking for lines that matched a given regular
expression, finally returning the total number of matched lines. Two series of
tests were run on two different files sizes. During these tests, we wanted to
see how Spark scales in an ideal setting, where the entire data set fit in
memory. This way we avoid the large performance hit Spark takes writing
intermediate results to disk. Given that each node has a total of 24 GB of RAM,
we limited our first problem size to 20GB. This allowed a single node to
contain the entire problem in memory, leaving sufficient memory space for the
OS and other supporting processes. We executed this first test with 1 to 5
nodes. The second test we performed was run with 5 to 50 nodes, each processing
a 100GB file.

@@

\emph{Data set}:
We obtained the data used for these tests from DBPedia, a crowd sourced
community that extracts structured information from Wikipedia and makes it
available on the web. \citep{[??]} There are different formats of Wikipedia
datasets available online. For our experiments we chose the Wikipedia pagelink
dataset \citep{[??]} , this contains internal links between the DBPedia
instances extracted from the internal links of the Wikipedia articles making it
apt for the any datamining related analysis including PageRank algorithm.
Therefore this dataset was chosen for the experiment.  The overall size of this
dataset is 25GB.  The format of the data is of the form: \\

\centering
\{ source url \} \{ some configuration \} \{ dest url \} \\

@@

\begin{figure}
\centering
  \includegraphics[width=80mm]{images/numVertices.png}
  \caption{The vertices count over various dataset size}
  \label{fig:numvertices}
\end{figure}

\begin{figure} 
  \centering
  \includegraphics[width=80mm]{images/numLinks.png}
  \caption{The average edge count per vertex for various dataset size}
  \label{fig:numlinks}
\end{figure}

The figure \ref{fig:numvertices} shows the variation in the no of vertices count for different dataset size which from ~40,000 nodes and the figure \ref{fig:numlinks} shows  the average number of links or edges per vertex for a smaller dataset is much higher as compared to the larger dataset. The smaller dataset are the first few DBPedia instances of the larger dataset. The first few pages probably have more links compared the pages in the end of the dataset and hence the average is linearly decreasing.

\emph{Algorithm:}
The basic idea of the page rank algorithm is to assign scores to the pages based on the following two aspects of the page.
	\begin{enumerate}
		\item {\centering pages with higher incoming links get high rank.}
		 \item {\centering pages with links from a high rank page also gets high rank.}
	\end{enumerate}

The pseudocode of the algorithm is as follows:
\begin{enumerate}
	\item {\centering Assign every page a rank of 1.}
	\item {\centering On each iteration have page p contribute to rankp/ |neighbours|p to its neighbors.}
	\item {\centering Set each page ranks to \(0.15 + 0.85  *  contributions\)}.
	\end{enumerate}

@@

In order to measure the relative performance of the Bagel and simple
implementations, we ran both versions of PageRank on
number of nodes from 4 to 16 with increment 2 and a dataset with 64,000 pages and
30 million links in total with sufficient memory. In Figure
\ref{fig:bagelsimple} regardless of time, we can see that bagel version
outperforms simple version a little bit. The main reason is that for a smaller
dataset, the overhead of managing vertices, edges and messages are so huge that
the benefits of asynchronous background messages that Spark might offer is not
that obvious. We couldn’t test on big dataset because of the aforementioned
technical failure. However, it's worth noting that programming is easier with
general concepts of vertices, edges and messages to represent the graph.

>>>>

\subsection{Testing}
Following the example from Google's first paper on MapReduce, the first test we
performed was a parallel grep \citep{dean-mapreduce}. Here an application would
scan through a large test file, looking for lines that matched a given regular
expression, finally returning the total number of matched lines. Two series of
tests were run on two different file sizes. During these tests, we wanted to
see how Spark scales in an ideal setting, where the entire data set fit in
memory. This way we avoid the large performance hit Spark takes writing
intermediate results to disk. Given that each node has a total of 24 GB of RAM,
we limited our first problem size to 20GB. This allowed a single node to
contain the entire problem in memory, leaving sufficient memory space for the
OS and other supporting processes. We executed this first test with 1 to 5
nodes. The second test we performed was run with 5 to 50 nodes, each processing
a 100GB file.

@@

\emph{Data set}:
We obtained the data used for these tests from DBPedia, a crowd sourced
community that extracts structured information from Wikipedia and makes it
available on the web. \citep{[??]} There are different formats of Wikipedia
datasets available online. For our experiments we chose the Wikipedia pagelink
dataset \citep{[??]} , this contains internal links between the DBPedia
instances extracted from the internal links of the Wikipedia articles making it
apt for the any datamining related analysis including PageRank algorithm.
Therefore this dataset was chosen for the experiment.  The overall size of this
dataset is 25GB.  The format of the data is of the form:

\begin{center}
\{ source url \} \{ some configuration \} \{ dest url \} \\
\end{center}

@@

\begin{figure}[H]
\centering
  \includegraphics[width=80mm]{images/numVertices.png}
  \caption{The vertices count over various dataset size}
  \label{fig:numvertices}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=80mm]{images/numLinks.png}
  \caption{The average edge count per vertex for various dataset size}
  \label{fig:numlinks}
\end{figure}

The figure \ref{fig:numvertices} shows the variation in the vertex count for
different dataset sizes which range from approximately 40,000 nodes and the
figure \ref{fig:numlinks} shows the average number of links or edges per vertex
for a smaller dataset is much higher as compared to the larger dataset. The
smaller dataset are the first few DBPedia instances of the larger dataset. The
first few pages probably have more links compared the pages in the end of the
dataset and hence the average is linearly decreasing.

\emph{Algorithm:}
The basic idea of the page rank algorithm is to assign scores to the pages
based on the following two aspects of the page.

\begin{enumerate}
    \item {\centering pages with higher incoming links get high rank.}
    \item {\centering pages with links from a high rank page also gets high
        rank.}
\end{enumerate}

The pseudocode of the algorithm is as follows:
\begin{enumerate}
    \item {\centering Assign every page a rank of 1.}
    \item {\centering On each iteration have page p contribute to $rankp/
        |neighbours|p$ to its neighbors.}
    \item {\centering Set each page ranks to \(0.15 + 0.85  *  contributions\)}.
\end{enumerate}

@@

In order to measure the relative performance of the Bagel and simple
implementations, we ran both versions of PageRank on
number of nodes from 4 to 16 with increment 2 and a dataset with 64,000 pages and
30 million links in total with sufficient memory. In Figure
\ref{fig:bagelsimple} regardless of time, we can see that bagel version
outperforms simple version a little bit. The main reason is that for a smaller
dataset, the overhead of managing vertices, edges and messages are so huge that
the benefits of asynchronous background messages that Spark might offer is not
that obvious. We couldn't test on a big dataset because of the aforementioned
technical failure. However, it's worth noting that programming is easier with
general concepts of vertices, edges and messages to represent the graph.
